---
title: "R Notebook"
output: html_notebook
---
# Install Packages
```{r}
#install.packages("tidyverse", dependencies = TRUE)
#install.packages("ggthemes", dependencies = TRUE)
```
# Load Packages
```{r}
library(tidyverse)
library(ggthemes)
library(rstatix)
library(ggpubr)
```
# Import Data
```{r}
assignment5 = read_csv("assignment5_data.csv")
```

# Subject Level Means
```{r}
subject_accuracy = assignment5 %>%
  group_by(subject, prime_condition) %>%
  summarise(mean_accuracy = mean(accuracy))

subject_rt = assignment5 %>%
  group_by(subject, prime_condition) %>%
  summarise(mean_RT = mean(response_RT))
```

# Format of the Data
Both subject_accuracy and subject_rt are in long format because they both have multiple rows for each observation (each subject has multiple rows for the different conditions).

# Long to Wide Format
```{r}
subject_accuracy_wide = subject_accuracy %>%
  pivot_wider(names_from = prime_condition, values_from = mean_accuracy)
```

# Wide to Long Format
```{r}
subject_accuracy_long = subject_accuracy_wide %>%
  pivot_longer(names_to = "prime condition", cols = both:unrelated)
```

# Interpretation 
The other data frame that includes this data is subject_accuracy because that was already in long format.

# T-Test in R
```{r}
t.test(subject_accuracy_wide$phonological, subject_accuracy_wide$semantic, var.equal = FALSE, paired = TRUE)
```

# T-Test interpretation
Because the t-test is 2-tailed, we compare our p value to 0.025 because that is half of alpha. Our p value of 9.33e-05 is much smaller than 0.025, which means there is a significant difference between the mean accuracy of the phonological and semantic conditions.

# T-Test Manual
```{r}
x_bar = mean(subject_accuracy_wide$mean_difference)
sd_diff = sd(subject_accuracy_wide$mean_difference)
t = x_bar/(sd_diff/sqrt(nrow(subject_accuracy_wide)))

df = nrow(subject_accuracy_wide) -1

p = 2*(1-pt(t,df))
```


# T-Test Outliers
```{r}
subject_accuracy_wide = subject_accuracy_wide %>%
  mutate(mean_difference = phonological - semantic)

subject_accuracy_wide = subject_accuracy_wide %>% ungroup()

hist(subject_accuracy_wide$mean_difference)

outliers = subject_accuracy_wide %>%
  identify_outliers(mean_difference)
```
There was one outlier, but it was not extreme so it does not need to be removed from the data.

# T Test Normality
```{r}
ggqqplot(subject_accuracy_wide, "mean_difference")

subject_accuracy_wide %>% shapiro_test(mean_difference)
```
The normality assumption is satisfied because p > .05, so there is no significance difference between our distribution and the normal distribution.

# Overall Pattern Interpretation
We have satisfied the assumptions about outliers and normality, so the t-test value is valid. This means that, on average, phonological cues help people to retrieve words more accurately than semantic cues.

# Plot RTs
```{r}
mean_rt = subject_rt %>%
  group_by(prime_condition) %>%
  summarise(mean_rt = mean(mean_RT))

mean_rt %>% 
  ggplot(aes(x = prime_condition, y = mean_rt)) +
  geom_col(fill = "pink", color = "navy")+
  theme_classic()+
  xlab("Prime Condition") + 
  ylab ("RT (in milliseconds)")+
  ggtitle("RT as a Function of Condition")
```
There was not much difference in RTs across prime condition. "Semantic" primes led to the slowest reaction times. Based on the plot, it is difficult to tell if "both" or "unrelated" primes led to faster reaction times, but according to the table, "both" primes led to the fastest reaction times.

# T-Test for RTs
```{r}
subject_rt_wide = subject_rt %>%
  pivot_wider(names_from = prime_condition, values_from = mean_RT)

subject_rt_wide = subject_rt_wide %>%
  mutate(mean_rt_diff = semantic - both)

subject_rt_wide = subject_rt_wide %>% ungroup()

rt_outliers = subject_rt_wide %>%
  identify_outliers(mean_rt_diff)

subject_rt_wide %>% shapiro_test(mean_rt_diff)

t.test(subject_rt_wide$semantic, subject_rt_wide$both, var.equal = FALSE, paired = TRUE)
```
If we again assume a two-tailed test, we compare our p value of 0.264 to 0.025 and find that there is no significant difference in the mean RT between the "semantic" and "both" prime conditions. 



